---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.
```{r}
library(pacman)
p_load(pastecs, crqa, ggplot2, stringr)
```

Define working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

```{r Load Data}

getwd()
locpath=getwd()
setwd(locpath)
```

1. In the course of this assignment you have to first select one datafile and figure out how to:
```{r}
pitch = read.table("New folder/Study1D0S101T3_f0.txt", header = TRUE)
```

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
```{r}
desc = stat.desc(pitch)
mean = desc[9,2]
sd = desc[13,2]
range = desc[6,2]

mean
sd
range
```

- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
```{r}
median = desc[8,2]
coefvar = desc[14,2]

median
coefvar
```

- Extract "complex" descriptors: recurrence quantification analysis
```{r}
# mutual(pitch$f0, lag.max = 50)
# #local minimum at 22
# 
# fnn =false.nearest(pitch$f0, m = 10, d = 22, t = 0)
# #how do I choose the values in this function?
# 
# plot(fnn)
# 
# ?mutual()
# ?false.nearest()

###

par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
ans = optimizeParam(pitch, pitch, par, min.rec = 3.5, max.rec = 4.5)

results = crqa(pitch, pitch, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

#represent the plot
RP = results$RP
RP = matrix(as.numeric(RP), nrow = ncol(RP))
cols = c("white", "blue4")
image(RP, xlab = "", ylab = "", col = cols)

#explore the lags of coordination
profile = drpdfromts(pitch, pitch, datatype = 'continuous', ws = 50, radius = ans$radius)
timecourse = round(seq(-5000, 5000, 100)/1000, digit = 1)
maxlag = profile$maxlag/1000
profile = profile$profile*100
prof = data.frame(profile)
ggplot(prof, aes(timecourse, profile)) + geom_line() + geom_vline(xintercept = timecourse[maxlag], colour = 'red')

#results
results
```


2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)
Study = Study
D0 = control
D1 = schizo
S = subject number/participant
T = trial
```{r}
files <- list.files(path = "C:/Users/Bruger/Dropbox/0Aarhus Universitet/Experimental Methods III/Assignment 3/ass3/Folderr")

setwd("C:/Users/Bruger/Dropbox/0Aarhus Universitet/Experimental Methods III/Assignment 3/ass3/Folderr")


meanPitch = NULL
SDPitch = NULL
crqaPitch = NULL
n = 1
participant = NULL
diagnosis = NULL
trial = NULL
PR = NULL
DET <- NULL
NRLINE <- NULL
maxL <- NULL
L <- NULL
ENTR <- NULL
LAM <- NULL
TT <- NULL
mean <- NULL
sd <- NULL
range <- NULL
median <- NULL
IQR = NULL
mad = NULL
coef_var = NULL
ans1 = NULL

for (i in files) {
  
  x = read.delim(i)
  x = x$f0
  
  ID = str_extract(i, "S+\\d+")
  participant[n] = str_extract(ID, "\\d+")
  Diagnosis = str_extract(i, "D+\\d")
  diagnosis[n] = str_extract(Diagnosis, "\\d")
  Trial = str_extract(i, "T+\\d+")
  trial[n] = str_extract(Trial, "\\d+")
  
  
  meanPitch[n] = mean(x, NA.RM = T)
  SDPitch[n] = sd(x)
  
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  ans1 = try(optimizeParam(x, x, par, min.rec = 3.5, max.rec = 4.5))
  
   if (length(ans1) > 1) {
  
  results1 <- crqa(x, x, delay = ans1$delay, embed = ans1$emddim, radius = ans1$radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  PR[n] = results1$RR
  DET[n] <- results1$DET
  NRLINE[n] <- results1$NRLINE
  maxL[n] <- results1$maxL
  L[n] <- results1$L
  ENTR[n] <- results1$ENTR
  LAM[n] <- results1$LAM
  TT[n] <- results1$TT
  
  }
  
  else {
  
  PR[n] = NA
  DET[n] <- NA
  NRLINE[n] <- NA
  maxL[n] <- NA
  L[n] <- NA
  ENTR[n] <- NA
  LAM[n] <- NA
  TT[n] <- NA
  
  } 
  
  n = n+1
}

df = data.frame(meanPitch, SDPitch, participant, diagnosis, trial, crqaPitch, PR, DET, NRLINE, maxL, L, ENTR, LAM, TT, mean, sd, range, median, IQR, mad, coef_var, ans1)
write.csv(df, file = "pitchfile.csv")

``` 

```{r}


#___________________
# make a list of the files that we are running through
files <- list.files(path = "C:/Users/Bruger/Dropbox/0Aarhus Universitet/Experimental Methods III/Assignment 3/ass3/New folder")
#files



# make empty list for all the information of interest
Subject <- NULL
study <-  NULL
trial <-  NULL
diagnosis <-  NULL
PR = NULL
DET <- NULL
NRLINE <- NULL
maxL <- NULL
L <- NULL
ENTR <- NULL
LAM <- NULL
TT <- NULL
mean <- NULL
sd <- NULL
range <- NULL
median <- NULL
IQR = NULL
mad = NULL
coef_var = NULL

# for every loop n will be increased by 1 so that we can distinguish between each file
n = 1

# for loop to run through each file in the list (files) we have already created above
for (file in files) {
  
  # read the file into a data frame to be used in the loop
  df = read.delim(file)
  
  # create the values to be put in each of the empty lists that we created above
  Subject[n] <-  substr(file, 14, 18)
  study[n] <-  substr(file, 7, 12)
  trial[n] <- sub("_", "", substr(file, 20, 21))
  diagnosis[n] <-  substr(file, 14, 14)
  mean[n] <- mean(df$f0)
  sd[n] <- sd(df$f0)
  range[n] <- range(df$f0)
  median[n] <- median(df$f0)
  IQR[n] <- IQR(df$f0)
  mad[n] <- mad(df$f0,center = mean(df$f0))
  coef_var[n] <- sd(df$f0)/mean(df$f0)*100
  
  # set parameterlist and optimize them
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  parameters <- try(optimizeParam(df$f0, df$f0, par, min.rec = 3.5, max.rec = 4.5))
  
  # this if/else statement assures that we the files that does not fit the parameters are still considered in the final lists 
  if (length(parameters) > 1) {
  
  results1 <- crqa(df$f0, df$f0, delay = parameters$delay, embed = parameters$emddim, radius = parameters$radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  PR[n] = results1$RR
  DET[n] <- results1$DET
  NRLINE[n] <- results1$NRLINE
  maxL[n] <- results1$maxL
  L[n] <- results1$L
  ENTR[n] <- results1$ENTR
  LAM[n] <- results1$LAM
  TT[n] <- results1$TT
  
  }
  
  else {
  
  PR[n] = NA
  DET[n] <- NA
  NRLINE[n] <- NA
  maxL[n] <- NA
  L[n] <- NA
  ENTR[n] <- NA
  LAM[n] <- NA
  TT[n] <- NA
  
  }
  
#increase n by 1 to index into the next value in the lists
  n = n + 1
  
}

# merge the list to a data frame
all_data_df_test <-  data.frame(Subject, study, trial, diagnosis, PR, DET, NRLINE, maxL, L, ENTR, LAM, TT, mean, sd, range, median)

#writedata frame to csv file
write.csv(all_data_df_test,file = "dfstudy1")
```


3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 
3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?

4. Bonus Question: Compare effect size of diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time